---
title: "Sky Island MRMs (permute modification)"
author: "Lili Benitez"
date: "most recent edits `r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: hide
    
---    
# Overview
Attempting to run MRMs on chao dissimilarity metrics for plants and polliantors across the Madrean Sky Islands. Code to modify the MRMs to account for site/year replication comes from Kaysee Arrowsmith with her permission. 

# Load Packages
```{r packages, message=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(geosphere)
library(ecodist)
library(Matrix)
library(lmtest)
library(tidyverse)
library(dplyr)
library(data.table)
library(ggplot2)
library(gdm)
library(CommEcol)
library(vegan)
library(conflicted)
theme_set(theme_classic())
```
# Import data
```{r}
setwd("C:/Users/lb37426/Box/Sky Islands Traits")
#netted bee community chao dissimilarity
bee.chao<-read.csv("net.bee.prob.csv")
rownames(bee.chao)<-bee.chao$X
bee.chao<-bee.chao%>%
  select(-X)
#flowering community chao dissimilarity data
flor<-read.csv("flor.chao.prob.csv")
rownames(flor)<-flor$X
flor<-flor%>%
  select(-X)
#precipitation 
precip<-read.csv("precip_4years.csv")%>%
  select(Precip, SiteYear)
rownames(precip)<-precip$SiteYear
precip<-precip%>%
  select(Precip)
precipitation_matrix<-as.matrix(dist(precip,method = "maximum",diag=TRUE))
#geographic distance
lat_long<-read.csv("unique.latlong.csv")
geo<-lat_long%>%
  select(Longitude, Latitude)
geo.dis <- distm(geo, fun = distGeo)
rownames(geo.dis)<-lat_long$SiteYear
colnames(geo.dis)<-lat_long$SiteYear
#convert to Km
geo.km<-geo.dis/1000# Apply distm function

#forest proportion
Forest<-read.csv("forest.prop.csv")%>%
  select(-Site)
rownames(Forest)<-Forest$SiteYear
Forest<-Forest%>%select(-SiteYear)
forest_dis<-as.matrix(dist(Forest,method = "maximum",diag=TRUE))
#write.csv(forest_dis, "Forest.dist.csv",row.names = FALSE)
```
#Here Kaysee slightly altered the `MRM` function from the `ecodist` package to accept a response variable in the form of `cbind(successes, failures)`. 

```{r mrm customize}
MRM.base <- function (formula = formula(data), data, nperm = 1000, method = "linear", mrank = FALSE, permute.control = "permute.control") {
  m <- match.call(expand.dots = FALSE)
  m2 <- match(c("formula", "data"), names(m), nomatch = 0)
  m <- m[c(1, m2)]
  m[[1]] <- as.name("model.frame")
  m <- eval(m, parent.frame())
  m <- as.matrix(m)
  n <- (1 + sqrt(1 + 8 * nrow(m)))/2
  if (abs(n - round(n)) > 1e-07) 
    stop("Matrix not square.\n")
  n <- round(n)
  if (ncol(m) < 2) 
    stop("Not enough data. \n")
  if (method == "linear") {
    X <- m[, 2:ncol(m), drop = FALSE]
      Y <- m[, 1, drop = FALSE]
      colnames(Y) <- "Y"
      newdata <- data.frame(Y = Y, X)
      fit1 <- lm(Y ~ ., data = newdata)
      b <- coefficients(fit1)
      sig <- summary(fit1)$sigma
      sig.df <- summary(fit1)$df[2]
      b.pval <- NA
      dev.pval <- NA
      if (nperm > 0) {
        b.all <- matrix(NA, nrow = nperm, ncol = length(b))
        b.all[1, ] <- b
        sig.all <- rep(NA, nperm)
        sig.all[1] <- sig
        for (i in 2:nperm) {
          newSample <- shuffle(n, control = permute.control)
          newY <- full(Y)
          newY <- newY[newSample, newSample]
          newY <- lower(newY)
          newdata <- data.frame(Y = newY, X = X)
          newfit <- lm(Y ~ ., data = newdata)
          b.all[i, ] <- coefficients(newfit)
          sig.all[i] <- summary(newfit)$sigma
        }
        b.pval <- apply(b.all, 2, function(x) length(x[abs(x) >= 
          abs(x[1])])/nperm)
        R2.pval <- length(sig.all[sig.all >= sig.all[1]])/nperm
      }
      results <- list(coef = cbind(b, pval = b.pval), 
        R2 = c(R2 = sig, R2.df = sig.df, 
          R2.pval = R2.pval))
  }
  else {
    if (method == "logistic") {
      X <- m[, 2:ncol(m), drop = FALSE]
      Y <- m[, 1, drop = FALSE]
      colnames(Y) <- "Y"
      newdata <- data.frame(Y = Y, X)
      fit1 <- glm(Y ~ ., data = newdata, family = binomial(link = "logit"))
      b <- coefficients(fit1)
      dev <- summary(fit1)$deviance
      dev.df <- summary(fit1)$df.residual
      b.pval <- NA
      dev.pval <- NA
      if (nperm > 0) {
        b.all <- matrix(NA, nrow = nperm, ncol = length(b))
        b.all[1, ] <- b
        dev.all <- rep(NA, nperm)
        dev.all[1] <- dev
        for (i in 2:nperm) {
          newSample <- shuffle(n, control = permute.control)
          newY <- full(Y)
          newY <- newY[newSample, newSample]
          newY <- lower(newY)
          newdata <- data.frame(Y = newY, X = X)
          newfit <- glm(Y ~ ., data = newdata, family = binomial(link = "logit"))
          b.all[i, ] <- coefficients(newfit)
          dev.all[i] <- summary(newfit)$deviance
        }
        b.pval <- apply(b.all, 2, function(x) length(x[abs(x) >= 
          abs(x[1])])/nperm)
        dev.pval <- length(dev.all[dev.all >= dev.all[1]])/nperm
      }
      results <- list(coef = cbind(b, pval = b.pval), 
        dev = c(resid.dev = dev, resid.df = dev.df, 
          dev.pval = dev.pval))
    }
    else {
      stop("method must be 'linear' or 'logistic'\n")
    }
    }
  results
}
```
## Analysis
```{r permutation control year}
# Permutation control for blocks and plots
siteyear.df <- data.frame(row = 1:nrow(bee.chao),
                        concat = rownames(bee.chao),
                        year = str_sub(rownames(bee.chao), start=-4), 
                        site=str_sub(rownames(bee.chao), start=1, end=2))


yearpermute.control <- how(Within(type = "free"),
                       blocks = siteyear.df$year)

sitepermute.control<-how(Within(type = "free"),
                       blocks = siteyear.df$site)
```
# test first with regular MRM
```{r}

#mrm
mrm_pols1<- suppressWarnings(MRM(as.dist(bee.chao)~as.dist(flor)+as.dist(precipitation_matrix)+as.dist(geo.km)+as.dist(forest_dis),
                method = "linear",
                 nperm = 1000,
                 ))


mrm_pols1

mrm_flor1<-suppressWarnings(MRM(as.dist(flor)~as.dist(precipitation_matrix)+as.dist(geo.km)+as.dist(forest_dis),
                method = "linear",
                 nperm = 1000,
                 ))
mrm_flor1

```
# Modified MRMs
```{r}
#Modified mrm with year permute control
mrm_pols<- suppressWarnings(MRM.base(as.dist(bee.chao)~as.dist(flor)+as.dist(precipitation_matrix)+as.dist(geo.km)+as.dist(forest_dis),
                method = "linear",
                 nperm = 1000,
                 permute.control = yearpermute.control))
mrm_pols

mrm_flor<-suppressWarnings(MRM.base(as.dist(flor)~as.dist(precipitation_matrix)+as.dist(geo.km)+as.dist(forest_dis),
                method = "linear",
                 nperm = 1000,
                 permute.control = yearpermute.control))
mrm_flor

# Modified mrm with site permute control
mrm_pols_2<- suppressWarnings(MRM.base(as.dist(bee.chao)~as.dist(flor)+as.dist(precipitation_matrix)+as.dist(geo.km)+as.dist(forest_dis),
                method = "linear",
                 nperm = 1000,
                 permute.control = sitepermute.control))
mrm_pols_2

mrm_flor_2<-suppressWarnings(MRM.base(as.dist(flor)~as.dist(precipitation_matrix)+as.dist(geo.km)+as.dist(forest_dis),
                method = "linear",
                 nperm = 1000,
                 permute.control = sitepermute.control))
mrm_flor_2

```

